---
title: "p8105_HW2_yj2686"
author: "Yiqun Jin"
date: "10/9/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

Read and clean the Mr. Trash Wheel sheet:

```{r Mr_Trash_Weel}
MrTrashData = 
  ##specify the sheet in the Excel file and to omit non-data entries
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
             sheet = "Mr. Trash Wheel", 
             range = "A2:N534" ) %>%
  ##use reasonable variable names
  janitor::clean_names() %>% 
  ##omit rows that do not include dumpster-specific data
  drop_na(dumpster) %>% 
  ## round the number of sports balls to the nearest integer
  mutate(sports_balls = round(sports_balls, 0))
```
Read and clean precipitation data for 2018 and 2019.

for 2018

```{r}
prcp_2018 = 
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
             sheet = "2018 Precipitation",
             range = "A2:B14") %>% 
  ##use reasonable variable names
  janitor::clean_names() %>% 
  ##omit rows without precipitation data
  drop_na() %>% 
  ##add a variable for year
  mutate(year = "2018") %>% 
  relocate(year)
```
for 2019

```{r}
prcp_2019 = 
  read_excel("./data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
             sheet = "2019 Precipitation",
             range = "A2:B14") %>% 
  ##use reasonable variable names
  janitor::clean_names() %>% 
  ##omit rows without precipitation data
  drop_na() %>% 
  ##add a variable for year
  mutate(year = "2019") %>% 
  relocate(year)
```
Combine precipitation datasets and convert month to a character variable

```{r}
combined_prcp =
  bind_rows(prcp_2018,prcp_2019) %>% 
  mutate(
    month = month.name[month]
  ) 
```

Summary of datasets

In Mr. Trash Wheel data, the key variables are `r names(MrTrashData)`. Mr. Trash Wheel data has `r nrow(MrTrashData)` observations of dumpsters, and `r ncol(MrTrashData) ` criteria to evaluate the trash in dumpsters. The median number of sports balls in a dumpster in 2019 is `r median(pull(filter(MrTrashData, year == 2019), sports_balls))`

In precipitation data, the key variables are `r names(combined_prcp)`. Precipitation data has `r nrow(combined_prcp)` number of observations in total. The total precipitation in 2018 is `r sum(select(filter(combined_prcp, year == 2018), total))`


## Problem 2

First, clean the data in pols-month.csv.

```{r}
pols_month = read_csv("./data/fivethirtyeight_datasets/pols-month.csv",show_col_types = FALSE) %>% 
  janitor::clean_names() %>% 
  ## break up the variable mon into integer variables year, month, and day
  separate(mon, into = c("year", "month","day")) %>%
  ## replace month number with month name
  mutate(month = as.integer(month),
         year = as.integer(year),
         month = month.name[month]) %>% 
  ## create a president variable taking values gop and dem
  mutate(president = if_else(prez_gop == 0, "dem", "gop")) %>%
  ## remove prez_dem, prez_gop, and day variables
  select(-prez_dem, -prez_gop, -day)
```

Second, clean the data in snp.csv.

```{r}
snp = read_csv("./data/fivethirtyeight_datasets/snp.csv", show_col_types = FALSE) %>% 
  janitor::clean_names() %>% 
  ## break up the variable date into integer variables year, month, and day
  mutate(date = lubridate::mdy(date)) %>% 
  separate(date, into = c("year", "month","day")) %>%
  ## replace month number with month name
  mutate(month = as.integer(month),
         month = month.name[month],
         day = as.integer(day),
         year = as.integer(year),
         year = ifelse(year > 2021, year - 100, year)) %>%
  select(-day) %>% 
  relocate(year, month) %>%
  arrange(year, month)
```

Third, tidy the unemployment data 

```{r}
unemploy = read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>%
  
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemployment") %>%
  
  janitor::clean_names() %>%

  mutate(month = month.name[match(month,month.abb)])

```


Join the datasets by merging snp into pols, and merging unemployment into the result.

```{r}
result = left_join(pols_month, snp)
result = left_join(result, unemploy)
```

Summary of datasets:

In pols_month dataset, the dimension of the dataset is `r dim(pols_month)`, the range of years is `r range(pull(pols_month, year))`, the key variables contains `r names(pols_month)`

In snp dataset, the dimension of the dataset is `r dim(snp)`, the range of years is `r range(pull(snp, year))`, the key variables contains `r names(snp)`

In unemployment dataset, the dimension of the dataset is `r dim(unemploy)`, the range of years is `r range(pull(unemploy, year))`, the key variables contains `r names(unemploy)`

After merging three datasets into result, the dimension of the dataset is `r dim(result)`, the range of years is `r range(pull(result, year))`, the key variables contains `r names(result)`

## Problem 3

Load and tidy the data. 

```{r}
popular_baby_names = 
  read.csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER"),
         ethnicity = recode(ethnicity, "WHITE NON HISP" = "WHITE NON HISPANIC"),
         ethnicity = recode(ethnicity, "BLACK NON HISP" = "BLACK NON HISPANIC"),
         child_s_first_name =  str_to_title(child_s_first_name),
         ) %>%
  distinct()
```

Produce a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time


```{r}
olivia_rank = 
  popular_baby_names %>% 
  filter(child_s_first_name == "Olivia", gender == "FEMALE") %>% 
  select(-count, -child_s_first_name,-gender) %>% 
  pivot_wider(
    names_from = year_of_birth,
    values_from = rank
  ) %>%
  knitr::kable(caption = "Rank in popularity of the name Olivia as a female baby name over time")
```

Produce a similar table showing the most popular name among male children over time.

```{r}
male_rank = 
  popular_baby_names %>%
  filter(gender == "MALE", rank == "1") %>% 
  select(-count, -rank, -gender) %>% 
  pivot_wider(
    names_from = year_of_birth,
    values_from = child_s_first_name
  ) %>%
  knitr::kable(caption = "Rank in popularity of the male baby name over time")
```

Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).

```{r}
male_white_non_hispanic_2016 = 
  popular_baby_names %>%
  filter(ethnicity == "WHITE NON HISPANIC", gender == "MALE", year_of_birth == "2016")

scatter_plot = 
  male_white_non_hispanic_2016 %>% 
  ggplot(aes(x = rank, y = count)) +
  geom_point() +
  labs(title = "Name rank of male white non-hispanic children born in 2016")
```





